Cell Decompositions for Rank Two Quiver Grassmannians

### 2018-01-17 Grassmannian_cells.tex
started improving section 5

\subsection{Lifting to $\widetilde{K(n)}$}

@@

The following result of Bialynicki-Birula relates the geometry of $X$ to the geometry of its torus fixed points.
\begin{theorem}
  \cite[Section 4]{bb}
  Let $\coprod_{i=1}^r C_i= X^{\C^\ast}$ be the decomposition into connected components.
  Then $\mathrm{Att}(C_i)$ is a locally closed smooth $\C^\ast$-invariant subvariety of $X$, where $C_i$ is a subvariety of $\mathrm{Att}(C_i)$.
  Moreover, we have $X=\coprod_{i=1}^r\mathrm{Att}(C_i)$ and the natural map $\gamma_i:\mathrm{Att}(C_i)\to C_i$ is an affine vector bundle.

@@

\subsection{Torus Actions}
\label{torusaction}

@@

\begin{theorem}
  Fix a dimension vector $\bfe$.
  The quiver Grassmannian $\Gr_\bfe(P_{m+1}^V)$ is smooth for each $V\in \Gr(H_m)$.
  Moreover, for $V,W\in \Gr_d(H_m)$, we have $\Gr_\bfe(P_{m+1}^V)\cong \Gr_\bfe(P_{m+1}^W)$.
\end{theorem}
\begin{proof}
  Let $g\in\GL_n(\CC)$.
  We first show that $\Gr_\bfe(g.P_{m+1}^V)=\Gr_\bfe(P_{m+1}^V)$.
  Assume that $P_{m+1}^V$ is given by the linear maps $M_{\alpha_i}$.
  Then $g.P_{m+1}^V$ is given by the matrices $(g.M)_{\alpha_i}=\sum\limits_{j=1}^n g_{ij}M_{\alpha_j}$.
  Let $(U_1,U_2)\in\Gr_\bfe(g.P_{m+1}^V)$.
  Then we have  
  \[M_{\alpha_i}(U_2)\subset U_1 \text{ for all } i=1,\ldots,n\Leftrightarrow \left(\sum\limits_{j=1}^n g_{ij}M_{\alpha_j}\right)(U_2)\subset U_1 \text{ for all } i=1,\ldots,n.\]
  Note that we have $g^{-1}.(g.P_{m+1}^V)=P_{m+1}^V$ which shows the non-obvious direction.

@@

\begin{theorem}\label{celldec}
  Every quiver Grassmannian of a preprojective or preinjective representation of $K(n)$ and $\widetilde{K(n)}$ has a cell decomposition.
\end{theorem}
\begin{proof}
This follows by combining the results of Section 2 and Proposition \ref{fibers}.
\end{proof}

@@

We recall the notion of Section \ref{QG} where we denote the arrows of $K(n)$ by $v_1,\ldots,v_n$.
Actually, all $2$-quivers can be built from the $2$-quivers of $P_1$, $P_2$ and $P_2^{(n-1)}$, where $P_2^{(n-1)}$ depends on a choice as it is a truncated preprojective representation of dimension $(1,1)$.
Indeed, it is the cokernel of a short exact sequence
\[
  \ses{ P_1^{n-1}}{ P_2}{P_2^{(n-1)}}
\]
for which we have $n$ choices.
Without loss of generality, we make the following choice when defining
\[
  P_2^{(n-1)}=j^{(n-1)}_2\xlongleftarrow{v_1}i^{(n-1)}_2=P_1^{(1)}=j^{(1)}_1\xlongleftarrow{v_1}i^{(1)}_1
\]
by its coefficient quiver.
Denoting by $j_1$ the single vertex of the coefficient quiver of $P_1$, we have 
\[
  \Ext(P_2^{(n-1)},P_1)=\langle i^{(n-1)}_{2}\xrightarrow{v_l}j_1\mid l=2,\ldots,n\rangle.
\]

>>>>

\subsection{Lifting to $\widetilde{K(n)}$}\label{Lifting}

@@

The following result of Bialynicki-Birula relates the geometry of $X$ to the geometry of its torus fixed points.
\begin{theorem}\label{thm:bb}
  \cite[Section 4]{bb}
  Let $\coprod_{i=1}^r C_i= X^{\C^\ast}$ be the decomposition into connected components.
  Then $\mathrm{Att}(C_i)$ is a locally closed smooth $\C^\ast$-invariant subvariety of $X$, where $C_i$ is a subvariety of $\mathrm{Att}(C_i)$.
  Moreover, we have $X=\coprod_{i=1}^r\mathrm{Att}(C_i)$ and the natural map $\gamma_i:\mathrm{Att}(C_i)\to C_i$ is an affine vector bundle.

@@

\subsection{Torus Actions}\label{torusaction}

@@

\begin{theorem}\label{thm:truncpp}
  Fix a dimension vector $\bfe$.
  The quiver Grassmannian $\Gr_\bfe(P_{m+1}^V)$ is smooth for each $V\in \Gr(H_m)$.
  Moreover, for $V,W\in \Gr_d(H_m)$, we have $\Gr_\bfe(P_{m+1}^V)\cong \Gr_\bfe(P_{m+1}^W)$.
\end{theorem}
\begin{proof}
  Let $g\in\GL_n(\CC)$.
  We first show that $\Gr_\bfe(g.P_{m+1}^V)=\Gr_\bfe(P_{m+1}^V)$.
  Assume that $P_{m+1}^V$ is given by the linear maps $M_{\alpha_i}$.
  Then $g.P_{m+1}^V$ is given by the matrices $(g.M)_{\alpha_i}=\sum\limits_{j=1}^n g_{ij}M_{\alpha_j}$.
  Let $(U_1,U_2)\in\Gr_\bfe(g.P_{m+1}^V)$.
  Then we have  
  \[M_{\alpha_i}(U_2)\subset U_1 \text{ for all } i=1,\ldots,n\Leftrightarrow \left(\sum\limits_{j=1}^n g_{ij}M_{\alpha_j}\right)(U_2)\subset U_1 \text{ for all } i=1,\ldots,n.\]
  Note that we have $g^{-1}.(g.P_{m+1}^V)=P_{m+1}^V$ which shows the non-obvious direction.

@@

\begin{theorem}\label{celldec}The following holds:
\begin{enumerate}
\item Every quiver Grassmannian of any indecomposable preprojective or preinjective representation of $K(n)$ and $\widetilde{K(n)}$ has a cell decomposition into affine spaces.
\item Let $X$ be an indecomposable representation of dimension $(d,e)$ or $(e,d)$ where $(d,e)=\udim P_m-r\udim P_{m-1}$ for $m\geq 2$ and $\leq 0\leq r\leq m-1$. Then every quiver Grassmannian $\Gr_{\tbfe}(X)$ has a cell decomposition into affine spaces.
\end{enumerate}
\end{theorem}
\begin{proof}
The first part follows by combining the results of Section \ref{torusaction} and Theorems \ref{thm:bb} and \ref{cellscover}. The dual version for preinjective representations follows as $\Gr_{\tbfe}(P_m)\cong\Gr_{\udim P_m-\tbfe}(I_m)$.

The second part follows in the same way, additionally, taking Theorem \ref{thm:truncpp} into account.
\end{proof}

@@

The next step is to associate a $2$-quiver with certain (truncated) preprojective representation $P_m^{\uj}=F_{K(n)}\tilde P_m^{\uj}$ for $\uj\in I_{n-1}$ whose strong successor closed subsets correspond to the cells of quiver Grassmannians. Therefore, we use the recursive construction described in Section \ref{Lifting}. In particular, we choose bases of the $\Ext$-groups which are compatible with the lifts to the universal covering. Recall that quiver Grassmannians $\Gr_{\tbfe}(P_m^V)$ attached to truncated preprojective representations only depend on $\tbfe$ and on $\udim P_m^V$.

Actually, all $2$-quivers can be built from the $2$-quivers of $P_1$, $P_2$ and $P_2^{(n-1)}$, where $P_2^{(n-1)}$ is chosen to be the cokernel of the short exact sequence
\[
  \ses{ P_1^{n-1}}{ P_2}{P_2^{(n-1)}.}
\]
Without loss of generality, we make the following choice when defining
\[
  P_2^{(n-1)}=j^{(n-1)}_2\xlongleftarrow{\alpha_1}i^{(n-1)}_2%=P_1^{(1)}=j^{(1)}_1\xlongleftarrow{v_1}i^{(1)}_1
\]
by its coefficient quiver.
Denoting by $j_1$ the single vertex of the coefficient quiver of $P_1$, we have 
\[
  \Ext(P_2^{(n-1)},P_1)=\langle i^{(n-1)}_{2}\xrightarrow{v_l}j_1\mid l=2,\ldots,n\rangle.
\]
